\section{Methodology} \locallabel{sec:methodology}
In the remainder of this chapter, we will use the term Hairball to refer to
both the Hairball framework and its set of plugins as described in
Section~\localref{sec:plugins}.

We tested Hairball on the \sprogram{s} submitted during our two-week summer
camp.  There were five assignments total, with a distribution of concept
requirements. For example, complex animation was taught toward the end of the
camp, thus instances of this concept were only present in the last two
assignments, whereas initialization was present in all
assignments~\cite{Franklin:2013:SBO}.

We first performed a manual analysis on all fifty-eight of the submitted
\sprogram{s}.  Three members of our staff independently analyzed the first five
\sprogram{s} submitted for a given assignment using a common rubric. We
discussed any discrepancies in our scores, and after coming to a consensus, we
analyzed the remaining \sprogram{s}. Once again, any score discrepancies were
reconciled.

Hairball was then programmed to match the methodology agreed upon by the staff
members when classifying the concepts, and subsequently used to statically
analyze all of the \sprogram{s}. We define our ground truth data set as all
instances that were labeled identically by both manual analysis and
Hairball. We performed a second manual analysis for each instance that Hairball
and the manual analysis labeled differently to determine which was correct,
Hairball or the initial manual analysis. The results of this second manual
analysis were added to our ground truth data set. In
Section~\localref{sec:results}, we compare both Hairball and our initial manual
analysis to our ground truth data set.

Because the assignments are sensory in nature (auditory, visual), we are not
attempting to create Hairball to replace manual analysis. Instead, we are
automating the identification of the \emph{easy} cases in order to accelerate
manual analysis of the remaining cases. As the results in
Section~\localref{sec:results} show, Hairball did an excellent job of
identifying issues that all three of our staff members missed.
