\begin{acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

The path to my Ph.D. was a long and tiresome one. I would never have reached
this point in my life without the love, influence, and support from other
people.

First and foremost, thank you Diana Franklin for taking me under your wing
these last few years. I could not have performed the research and completed
this dissertation without your patience and guidance.

The loss of my brother nearly resulted in me abandoning the completion of my
Ph.D. I would have done so without the generosity of Klaus Schauser who
mentored me and provided me funding through Appfolio, Inc. during the final
stage of my Ph.D. Thank you Klaus, and thank you Appfolio, Inc.

Mom, thank you for providing me with the freedom to learn. Dad, thank you for
showing me how the world works. Collectively, you influenced my development of
both an inquisitive nature and a desire to learn. Furthermore, thank you for
tolerating me breaking the computer on a plethora of occasions.

Matt, Daniel (1988--2014), and Connor, thank you for being the best brothers
anyone could ask for. Buz, thank you for introducing me to HTML in \nth{6}
grade, and being a great friend through all these years. Adam and Scott, thank
you for the camaraderie, support, and encouragement during our computer science
education. To all the gentlemen of the GALAGA house, both past and present,
thank you for all the great events and outings. I am eternally grateful to have
met such amazing friends.

Tim, thank you for organizing the 2004 programming battle and subsequently
mentoring me throughout my college education. Andy, thank you for the
encouragement and all the challenging tasks you entrusted to me at WorldViz,
LLC. Cha and Bob, thank you for convincing me to apply to the Ph.D. program by
attempting to discourage me from doing so. Janet and Peter, and Cheri and Tom,
thank you for the unrelenting support and confidence in my success. Thank you
to all my past teachers who challenged me and encouraged me to succeed.

Sharon and George, thank you for the love, the support, and all the delicious
home cooked meals over the last year. My new sisters, brothers, nieces,
nephews, aunts, uncles, and cousins, thank you making me feel welcome, for the
support when my brother died, and for the encouragement to complete this
Ph.D. adventure.

Finally, Julie, my love, thank you for the incredible support during the final
year of my Ph.D. I would not have made it through this last year without your
patience, understanding, and love. The next chapter of my life belongs to you.

\end{acknowledgements}
