\section{Conclusion} \locallabel{sec:conclude}
In this study we discover student submission behaviors through the analysis of
20,777 submissions made by 289 students across seven classes. The data was
collected by a real-time feedback system we created that allowed a
per-assignment feedback delay to be configured. Our results show that delaying
feedback impacts student submission behavior. Furthermore, delays of at least
thirty minutes positively affects the efficiency of submissions when compared
to smaller delays. Our results also suggest that delaying feedback impacts
student work sessions in two ways:

\begin{itemize}
\item Increases in delay correlate with longer work sessions
  (Section~\localref{sec:efficiency}).
\item Increases in delay correlate with less improvement during work sessions
  (Section~\localref{sec:session_impact}).
\end{itemize}

The aggregate results of the feedback delay suggests that assignments should be
configured with a thirty minute feedback delay. Our results also provide an
interesting comparison to prior work:

\begin{itemize}
\item We confirm that starting early correlates with higher assignment
  scores (Section~\localref{sec:early}).
\item We confirm a high-period (not peak) of student activity between \PM{4}
  and \PM{6} regardless of deadline (Section~\localref{sec:time}).
\item We identify peak student activity occurring in the hours prior to the
  time of a deadline both on the day of the deadline and others. Due to our
  differences with prior work, we hypothesize that students adjust their peak
  working hours to align with pending deadlines (Section~\localref{sec:time}).
\item We confirm that a majority of activity is completed in the two days prior
  to assignment deadline (Section~\localref{sec:efficiency}).
\end{itemize}

In addition to our comparison with prior work we offer some other new insights:

\begin{itemize}
\item We show that there is no difference in submission efficiency with
  submission time relative to assignment deadline
  (Section~\localref{sec:efficiency}).
\item We discuss three reasons why students submit well after the deadline
  (Section~\localref{sec:deadline}).
\item We describe a method for window size selection that balances the two
  forms of error (Section~\localref{sec:window}).
\end{itemize}

Overall, this study provided an insight into a few aspects of student
submission behavior. While more research is required to fully understand
student submission behavior, our results should help guide instructors towards
assignment configuration with respect to feedback delay and assignment
deadlines.
